mruby Bytecode Loader Using Bluetooth in Multi-VM Environment

### 2016-02-11 thesis/thesis.tex
update

%
% File: thesis.tex
% Created: 月 11 30 08:00 午後 2015 東京 (標準時)
% Last Change: 月 11 30 08:00 午後 2015 東京 (標準時)
%
\input{dummy/dummy.tex}
\documentclass[conference,compsoc]{IEEEtran}
\bibliographystyle{IEEEtranBST2/IEEEtran}
\usepackage{url}
\usepackage[dvipdfmx]{graphicx}
\usepackage{amssymb}
\usepackage{listings,jlisting}
\lstset{%
    language={},
    basicstyle={\small},%
    identifierstyle={\small},%
    commentstyle={\small\itshape},%
    keywordstyle={\small\bfseries},%
    ndkeywordstyle={\small},%
    stringstyle={\small\ttfamily},
    frame={tb},
    breaklines=true,
    columns=[l]{fullflexible},%
    numbers=left,%
    xrightmargin=0zw,%
    xleftmargin=3zw,%
    numberstyle={\scriptsize},%
    stepnumber=1,
    numbersep=1zw,%
    lineskip=-0.5ex%
}

\title{mruby Bytecode Loader Using Bluetooth in Multi-VM Environment}
\author{Takuro Yamamoto}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
In recent years, the productivity of embedded systems has become a problem due to their complexity and large-scale.
For the purpose of improving the productivity for embedded software development, the mruby on TECS framework has been proposed that is applied mruby (Lightweight Ruby) and supports component-based development.
In the current mruby on TECS, the mruby programs have to be compiled and linked every time the programs are modified because the mruby bytecodes are incorporated in the platform.
Moreover, while the framework supports multi-VM, developers need to be familiar with the functions of real-time operating systems to effectively execute multiple mruby programs in concurrent or/and parallel.
To improve the development efficiency, this paper proposes an mruby bytecode loader using Bluetooth as an extension of mruby on TECS.
The loader executes two mruby bytecodes, mruby application bytecode and mruby library bytecode.
mruby application bytecode modified frequently is sent from a host to a target device by developers.
mruby library bytecode modified infrequently is preserved beforehand in an SD card at the time of the first compailation.
In addition, multiple mruby programs cooperatively run in the proposed framework.
In the proposed framework, the cyclic handler switching tasks is implemented, and multitasking processing is more easy-to-use than that of mruby on TECS.
Experimental results demonstrate the advantages of the proposed framework.
\end{abstract}
\section{Introduction}
In these days, embedded systems have been required the high-quality and the high-performance.
Due to this trend, the complexity of embedded systems also increases and the scale is larger.
For example, IoT (Internet of Things) applications.
The low production cost and the short developing period of time are also needed.

@@

Although script languages are easy to use and read, their execution time are slower than C language's.
For embedded systems, the real-time properties such as worst-execution time, response time are very important factors.
Therefore, it is difficult to apply the script languages to embedded systems.

mruby on TECS is a component-based framework for running script program \cite{par:mrubyonTECS}.
It is integrated two technologies.
One is mruby, which is a script language for embedded systems \cite{par:mruby}, \cite{url:mruby}.
The other is TECS (TOPPERS Embedded Component System), which is a component-based framework for embedded systems \cite{par:TECS}.
TECS supports to effectively run mruby script language on embedded systems.
mruby on TECS also makes execution time 100 times faster than that of mruby.

This paper proposes two additional features of mruby on TECS, mruby bytecode loader using Bluetooth and user-friendly multitask.
mruby on TECS has some problems at present.
One of the problems is that working efficiency is low because mruby on TECS only supports an SD card in the platform for LEGO MINDSTORMS EV3 \cite{par:EV3} to load mruby programs.
In the current development process, it is annoying that the SD card should be insert and pulled out repeatedly.
If a new feature, mruby bytecode loader Bluetooth, is used, the procedure is carried out only once at the beginning and a program code is transferred from a host to a target device by Bluetooth.
It can improve the efficiency.
Moreover, in the current system if developers attempt to run multiple tasks with the same priority, a task with first execution only runs and all other tasks would not run unless a OS's function must be called by developers.
In addition developers must prepare as many VMs as the number of threads because one RiteVM supports only one thread.
Thus that is a big burden on developers.
Another new feature enables developers to run a multitask programs more easily.
Specifically, a cyclic handler of switching tasks is implemented.
This paper evaluates the proposed framework in term of data transfer rate, overhead of cyclic handler and execution time.

{\bf Contributions}: This paper proposes mruby bytecode loader using Bluetooth and user-friendly multitasking.
The proposed framework gives the contribution in the following points.
\begin{enumerate}
\item To improve the  work efficiency in software development on mruby on TECS: \mbox{}\\
\ mruby bytecode loader using Bluetooth helps developers develop software.
Developers do not need to insert and pull out an SD card repeatedly and also to restart an OS.
Thus, the software development efficiency is improved.
\item To effectively execute multiple mruby programs in concurrent or/and parallel: \mbox{}\\
\ There is a problem in multitasking of the current mruby on TECS.
It is that developers need to be well acquainted the RTOS if multiple mruby programs execute in multitasking.
In the proposed framework, it is easy for developers to use multitasking because multiple tasks are switched cyclically.
Multitasked programing makes the range of software expand.
\end{enumerate}

{\bf Organization}: The paper is organized as follows.
Section \ref{sec:Background} introduces the basic technologies i.e. mruby, TECS and mruby on TECS.
Section \ref{sec:mruby bytecode loader using Bluetooth} and Section \ref{sec:Multitask} describe the design and implementation of the proposed framework in detail.
Section \ref{sec:Evaluation} evaluates the proposed framework, Section \ref{sec:Related work} discuses related work, and then Section \ref{sec:Conclusion} concludes.

\section{Background}
\label{sec:Background}
Figure \ref{fig:proposed} shows a system model of the proposed framework.
The bytecodes are transferred from the host to the target device with Bluetooth.
The RiteVMs and mruby library are assumed to be prepared in advance.
Each bytecode is allocated to a RiteVM, respectively.
Developers can run some bytecodes transferred with Bluetooth in multitask.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/proposed.pdf}
    \caption{System Model}
    \label{fig:proposed}
\end{figure}

@@

\subsection{mruby}
\label{sec:mruby}
mruby is the light-weight implementation of Ruby programming language complying to part of the ISO standard.

@@

mruby is suitable for embedded systems because of faster execution with less amount of resources and takes over the usability and readability of Ruby.
In addition, VM (Virtual Machine) mechanism is used in mruby, so mruby programs can run on any operating system as long as VM is implemented.
RiteVM is a VM in mruby, that runs mruby programs.
The RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which is an intermediate code that can be interpreted by RiteVM.
The bytecodes can run on a Rite VM, and thus mruby programs can be executed on any target devices if only RiteVMs are implemented.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby.pdf}
    \caption{Mechanism of mruby/RiteVM}
    \label{fig:mruby}
\end{figure}

\subsection{TECS}
\label{sec:TECS}
TECS (TOPPERS Embedded Component System) is a component system suitable for embedded systems.
TECS helps decrease the complexity and difficulty because the generated component diagram can visualize the structure of whole software.
It can also help increase the productivity and reduce development duplication  because a common part of software is regarded as a component.

@@

\subsubsection{Component Model}\mbox{}\\

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_diagram.pdf}
    \caption{Component Diagram}
    \label{fig:component}
\end{figure}

\subsubsection{Component Description}\mbox{}\\

@@

\subsection{mruby on TECS}
\label{sec:mruby on TECS}
mruby on TECS is a component-based framework for running script language.
This framework uses two technologies, mruby and TECS.

\subsubsection{System Model}\mbox{}\\

The system model of mruby on TECS is shown in Figure \ref{fig:system_model}.
Each mruby program, which is bytecode, runs on its own RiteVM as a componentized task of an RTOS.
TECS components support various embedded drivers such as motor and sensor drivers.

An mruby-TECS bridge plays a role to call a native program (e.g.,C legacy code) from an mruby program.
The mruby-TECS bridge provides native libraries for mruby.
It also gives TECS components to receive the invocation from an mruby program.
The mruby-TECS bridge is described in more detail bellow.

As the target RTOS, TOPPERS/HRP2 \cite{url:HRP2}, \cite{par:hr-tecs} is used in this paper.
TOPPERS/HRP2 is an RTOS based on $\mu$ITRON with memory protection.
However, mruby on TECS does not depend on the RTOS because TECS supports not only TOPPERS/HRP2 but also the other RTOSs such as OSEK \cite{par:OSEK} and TOPPERS/ASP \cite{url:ASP}.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/system_model.pdf}
    \caption{System Model}
    \label{fig:system_model}
\end{figure}

\subsubsection{mruby-TECS Bridge}\mbox{}\\

@@

Figure \ref{fig:mruby_TECS_bridge} shows an example of use of an mruby-TECS bridge for controlling a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby_TECS_bridge.pdf}
    \caption{mruby-TECS Bridge}
    \label{fig:mruby_TECS_bridge}
\end{figure}

@@

 \section{mruby Bytecode Loader Using Bluetooth}
\label{sec:mruby bytecode loader using Bluetooth}
This section describes an additional feature of mruby on TECS, mruby bytecode loader using Bluetooth.
In the current system, all binary data including bytecodes are in an SD card.
Developers must insert and pull out the SD card every time the source programs are modified.
It causes a low work efficiency that such a work must be repeated.
mruby bytecode loader using Bluetooth makes the developers' burden decrease.
The work that an SD card is in and out needs to be done just once in the beginning.

mruby programs are consisted of mruby application and libraries.
mruby application is the main code of the application.
mruby libraries are the code defining the functions for application such as Ruby class. 
The mruby bytecodes including mruby application and libraries can be sent and run.
However it is wasteful in terms of the size and time to send because libraries are not frequently modified. 
In the proposed framework, mruby application only sends, and mruby libraries are preserved in the SD card beforehand.

The development flow in mruby bytecode loader using Bluetooth is shown in Figure \ref{fig:bluetooth_loader}.
The RiteVM and mruby library are compiled and copied in the SD card at the first.
These are assumed not to be modified.
The binary data transferred with Bluetooth is the bytecode of the main source.
In the host, the source files (.rb) are edited and compiled into the bytecodes (.mrb) by an mruby compiler.
The generated bytecodes are transferred from the host to the target device with Bluetooth.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/bluetooth_loader.pdf}
    \caption{Development Flow in mruby bytecode loader using Bluetooth}
    \label{fig:bluetooth_loader}
\end{figure}

\subsection{Component of mruby bytecode loader using Bluetooth}
The proposed framework provides components of mruby bytecode loader using Bluetooth to use the function.
A component of mruby bytecode loader using Bluetooth is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages RiteVM configuration such as automatically generates the bytecode in the build description.
This generated bytecode is prepared beforehand in the SD card such as mruby libraries, and different from a bytecode transferred with Bluetooth.

Figure \ref{fig:component_bluetooth} shows a component diagram of MrubyTask1 and MrubyBluetooth1 {\it cell}s.
The MrubyTask1 {\it cell} is a componentized task of the RTOS (TOPPERS/HRP2).
TOPPERS/HRP2 is described in \cite{url:HRP2}, \cite{par:ht-tecs}.
The MrubyBluetooth1 is a component of mruby bytecode loader using Bluetooth.
A bytecode which is a compiled mruby program in the host is transferred and received at the top of the mruby bytecode loader using Bluetooth component.
In this framework, ZMODEM \cite{par:zmodem} is used as a binary transfer protocol.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_bluetooth.pdf}
    \caption{Component Diagram of mruby bytecode loader using Bluetooth}
    \label{fig:component_bluetooth}
\end{figure}

Figure \ref{fig:control_flow} shows the process of executing mruby program in a component of mruby bytecode loader using Bluetooth, which is like MrubyBluetooth1.
First, a pointer of {\it mrb\_state} structure is initialized.
{\it mrb\_state} is a set of states and global variables used in mruby.
Second, the bytecode of mruby libraries is read.
mruby libraries are a set of Ruby classes such as motor class and sensor class.
For example, motor class defines methods to rotate and stop a motor.
The tMrubyBluetooth {\it cell} has two attributes as shown in Figure \ref{celltype_mrubybluetooth}: {\it mrubyFile} and {\it irep}.
The {\it mrubyFile} indicates the program files of mruby libraries.
{\it [omit]} is only used for the TECS generator, thus the attribute, {\it mrubyFile}, does not consume memory.
The irep is the pointer of the array stored the bytecode of mruby libraries.
In short the bytecode of mruby libraries is stored as an attribute of the component when compiling for the first time.
Third, the bytecode of the mruby program transferred with Bluetooth is read .
The bytecode transmitted via Bluetooth is stored in an array of type uint8\_t, which is the same as type char.
The array is different from that of holding the mruby library bytecode.
Two bytecodes are read separately in the {\it mrb\_state}.
Finally, an mruby task runs.
When the mruby program is modified, only the bytecode of the modified program should be transferred.
mruby libraries need not be touched because libraries are not normally changed.
The concrete main code of tMrubyBluetooth is shown in Figure \ref{maincode_mrubybluetooth}.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/control_flow.pdf}
    \caption{Control Flow of mruby bytecode loader using Bluetooth}
    \label{fig:control_flow}
\end{figure}
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tMrubyBluetooth{
    entry sTaskBody eMrubyBody;
    attr{
      [omit]char_t *mrubyFile;
      char_t *irepLib=C_EXP("&$cell_global$_irep");
      uint8_t irepApp[COUNT_STK_T(TMAX_APP_BINARY_SIZE)]
              __attribute__((nocommon));
    };
};
\end{lstlisting}
\caption{Celltype Description for mruby bytecode loader using Bluetooth}
\label{celltype_mrubybluetooth}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
void
eMrubyBody_main(CELLIDX idx)
{
  /* Declaration variables */
  mrb_state *mrb;
  mrbc_context *c;
  /* New interpreter instance */
  mrb = mrb_open();
  // Omit: error check for mrb_state
  /* New mruby context */
  c = mrbc_context_new(mrb);
  // Omit: initialization of mruby-TECS bridge
  /* Receive the bytecode via Bluetooth */
  bluetooth_loader( ATTR_irepApp );
  /* Load mruby library bytecode and run */
  mrb_load_irep_cxt(mrb, ATTR_irepLib, c);
  /* Load mruby application bytecode and run */
  mrb_load_irep_cxt(mrb, ATTR_irepApp, c);
  if (mrb->exc) {
    /* Failure to execute */
    mrb_p(mrb, mrb_obj_value(mrb->exc));
    exit(0);
  }
  /* Free mruby context */
  mrbc_context_free(mrb, c);
  /* Free interpreter instance */
  mrb_close(mrb);
}

@@

One of approaches for multitasking is co-routine.
Co-routine is a cooperative thread, and scheduled by developers with the functions such as {\it resume} and {\it yield}. 
(Ruby co-routine is defined in class Fiber \cite{url:co-routine})
Co-routine is a non-preemptive mulititasking, and does not receive the OS's support because developers have to switch tasks manually.
Co-routine can not take advantege of multi core processing.

Besides, as another method, {\it delay}, a service call of $\mu$ITRON, can be used for multitasking.
This service call delays the execution of the own task for the time of the argument.
In the approach of {\it delay} fubction, developers have to schedule as the same as co-routine.

In the proposed framework, as an approach for multitask processing, {\it rotateReadyQueue}, a service call of $\mu$ITRON, is implemented.
{\it rotateReadyQueue} rotates the precedence of the task with the priority of the argument.
A cyclic handler calls {\it rotateReadyQueue}, which enables developers to develop applications without concern for multitasking.

>>>>

%
% File: thesis.tex
% Created: 月 11 30 08:00 午後 2015 東京 (標準時)
% Last Change: 月 11 30 08:00 午後 2015 東京 (標準時)
%
\input{../EMSOFT2016/dummy/dummy.tex}
\documentclass[a4j,12pt,oneside,openany,english]{jsbook}
\usepackage{url}
\usepackage[dvipdfmx]{graphicx}
\usepackage{amssymb}
\usepackage{utf8/book2report}
\usepackage{utf8/thesis}
\usepackage{lscape}
\usepackage{listings,jlisting}
\lstset{%
    language={},
    basicstyle={\small},%
    identifierstyle={\small},%
    commentstyle={\small\itshape},%
    keywordstyle={\small\bfseries},%
    ndkeywordstyle={\small},%
    stringstyle={\small\ttfamily},
    frame={tb},
    breaklines=true,
    columns=[l]{fullflexible},%
    numbers=left,%
    xrightmargin=0zw,%
    xleftmargin=3zw,%
    numberstyle={\scriptsize},%
    stepnumber=1,
    numbersep=1zw,%
    lineskip=-0.5ex%
}

% \title{mruby Bytecode Loader Using Bluetooth in Multi-VM Environment}
% \author{Takuro Yamamoto}
% \date{\today}

\begin{document}
\thispagestyle{empty}
\begin{center}
\vspace*{5mm}
{\Huge {\bf Bachelor's Thesis}}\\
\vspace{2cm}
%{\Large 題\hspace{8mm}目}\\
%\vspace{1cm}
\underline{\LARGE{mruby Bytecode Loader}} \\
\vspace{0.5cm}
\underline{\Large{Using Bluetooth in Multi-VM Environment}} \\
\vspace{12mm}
{\large Supervisor}\\
\vspace{6mm}
\underline{\Large Professor Toshimitsu Ushio}\\
\underline{\Large Assistant Professor Takuya Azumi}\\
\vspace{8mm}
{\large By}\\
\vspace{6mm}
\underline{\Large Takuro Yamamoto}\\
\vspace{10mm}
{\Large February 15, 2016}\\
\vspace{14mm}
{\Large Division of Systems Science and Applied Informatics,\\Department of Systems Science,\\Schoolo of Engineering Science, Osaka University}\\
\end{center}
\clearpage
\setcounter{page}{0}
\pagenumbering{roman}
%\maketitle

\begin{abstract}
In recent years, the productivity of embedded systems has become a problem due to their complexity and large-scale.
For the purpose of improving the productivity for embedded software development, the mruby on TECS framework has been proposed that is applied mruby (Lightweight Ruby) and supports component-based development.
In the current mruby on TECS, the mruby programs have to be compiled and linked every time the programs are modified because the mruby bytecodes are incorporated in the platform.
Moreover, while the framework supports multi-VM, developers need to be familiar with the functions of real-time operating systems to effectively execute multiple mruby programs in concurrent or/and parallel.
To improve the development efficiency, this paper proposes an mruby bytecode loader using Bluetooth as an extension of mruby on TECS.
The loader executes two mruby bytecodes, mruby application bytecode and mruby library bytecode.
mruby application bytecode modified frequently is sent from a host to a target device by developers.
mruby library bytecode modified infrequently is preserved beforehand in an SD card at the time of the first compilation.
In addition, multiple mruby programs cooperatively run in the proposed framework.
In the proposed framework, RiteVM sceduler makes multitasking processing more easy-to-use than that of mruby on TECS.
Synchronization of starting multiple tasks is also implemented with Eventflag. 
Experimental results demonstrate the advantages of the proposed framework.
\end{abstract}
\clearpage
\tableofcontents
\clearpage
\setcounter{page}{0}
\pagenumbering{arabic}
\chapter{Introduction}
In these days, embedded systems have been required the high-quality and the high-performance.
Due to this trend, the complexity of embedded systems also increases and the scale is larger.
For example, IoT (Internet of Things) applications.
The low production cost and the short developing period of time are also needed.

@@

Although script languages are easy to use and read, their execution time are slower than C language's.
For embedded systems, the real-time properties such as worst-execution time and response time are very important factors.
Therefore, it is difficult to apply the script languages to embedded systems.

mruby on TECS is a component-based framework for running script program \cite{par:mrubyonTECS}.
It is integrated two technologies.
One is mruby, which is a script language for embedded systems \cite{par:mruby}, \cite{url:mruby}.
The other is TECS (TOPPERS Embedded Component System), which is a component-based framework for embedded systems \cite{par:TECS} \cite{url:TOPPERS}.
mruby on TECS supports to effectively run mruby script language on embedded systems.
mruby on TECS also makes execution time 100 times faster than that of mruby.

This paper proposes two additional features of mruby on TECS, mruby bytecode loader using Bluetooth and user-friendly multitask.
mruby on TECS has several problems at present.
One of the problems is that working efficiency is low because mruby on TECS only supports an SD card in the platform for LEGO MINDSTORMS EV3 \cite{par:EV3} to load mruby programs.
In the current development process, it is annoying that the SD card should be inserted and pulled out repeatedly and the OS should be restarted.
%Moreover, in the current system if developers attempt to run multiple tasks with the same priority, a task with first execution only runs and all other tasks do not run unless developers call the OS's function.
%In addition developers must prepare as many VMs as the number of tasks because one RiteVM supports only one task.
Although mruby on TECS has supported the multi-VM, developers need to call OS's function in order to execute multiple tasks.
Thus, that is a big burden on developers.

This paper proposes mruby bytecode loader using Bluetooth and RiteVM scheduler for fairly executing mruby programs.
To improve the development efficiency, mruby bytecode loader using Bluetooth enables developers to insert an SD card only once at the beginning and to transfer mruby application codes from a host to a target device.
RiteVM scheduler enables developers to run multitasking programs more easily than mruby on TECS.
%Specifically, a cyclic handler of switching tasks is implemented.
%This paper evaluates the proposed framework in term of data transfer rate, overhead of cyclic handler and execution time.

{\bf Contributions}: The proposed framework gives the contribution in the following points.
\begin{enumerate}
\item To improve the  work efficiency in software development on mruby on TECS: \mbox{}\\
\ mruby bytecode loader using Bluetooth helps developers develop software.
Developers do not need to insert and pull out an SD card repeatedly and also to restart an OS.
The loader supports the countinuous loading, which saves the pairing time for Bluetooth.

\item To effectively execute multiple mruby programs in concurrent or/and parallel: \mbox{}\\
\ There is a problem in multitasking of the current mruby on TECS.
It is that developers need to be well acquainted with the RTOS if multiple mruby programs execute in multitasking.
In the proposed framework, it is easy for developers to use multitasking because RiteVM sceduler switches tasks cyclically.
Synchronization of multipe tasks is also implemented.

\item To focus on the advantage of component-based developments: \mbox{}\\
\ The proposed framework has been developed with component-based.
Component-based development gives various benefits.
 This paper shows the specific examples for the benefits of component-based development. 
\end{enumerate}

{\bf Organization}: The paper is organized as follows.
Section \ref{sec:Background} introduces the basic technologies i.e. mruby, TECS and mruby on TECS.
Section \ref{sec:Design and Implementation} describes the design and implementation of the proposed framework in detail.
Section \ref{sec:Evaluation} evaluates the proposed framework, Section \ref{sec:Related work} discusses related work, and then Section \ref{sec:Conclusion} concludes this paper.

\chapter{Background}
\label{sec:Background}
Figure \ref{fig:proposed} shows the system model of the proposed framework.
The bytecodes are transferred from the host to the target device with Bluetooth.
The RiteVMs and mruby library are assumed to be prepared in advance.
Each bytecode is allocated to a RiteVM, respectively.
Developers can run bytecodes transferred with Bluetooth in multitask.

\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/proposed.pdf}
    \caption{System Model}
    \label{fig:proposed}
\end{figure}

@@

\section{mruby}
\label{sec:mruby}
mruby is the light-weight implementation of Ruby programming language complying to part of the ISO standard.

@@

mruby is suitable for embedded systems because of faster execution with less amount of resources and takes over the usability and readability of Ruby.
In addition, VM (Virtual Machine) mechanism is used in mruby, therefore mruby programs can run on any operating system as long as VM is implemented.
RiteVM is a VM in mruby, that runs mruby programs.
The RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which is an intermediate code that can be interpreted by RiteVM.
The bytecodes can run on a Rite VM, and thus mruby programs can be executed on any target devices if only RiteVMs are implemented.
\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/mruby.pdf}
    \caption{Mechanism of mruby/RiteVM}
    \label{fig:mruby}
\end{figure}

\section{TECS}
\label{sec:TECS}
TECS (TOPPERS Embedded Component System) is a component system suitable for embedded systems.
TECS helps decrease complexity and difficulty because the generated component diagram can visualize the structure of whole software.
It can also help increase productivity and reduce development duplication  because a common part of software is regarded as a component.

@@

\subsection{Component Model}\mbox{}\\

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/component_diagram.pdf}
    \caption{Component Diagram}
    \label{fig:component}
\end{figure}

\subsection{Component Description}\mbox{}\\

@@

\section{mruby on TECS}
\label{sec:mruby on TECS}
mruby on TECS is a component-based framework for running script language.
This framework uses two technologies, mruby and TECS.

\subsection{System Model}\mbox{}\\

The system model of mruby on TECS is shown in Figure \ref{fig:mrubyontecs}.
Each mruby program, which is bytecode, runs on its own RiteVM as a componentized task of an RTOS.
TECS components support various embedded drivers such as motor and sensor drivers.

An mruby-TECS bridge plays a role to call a native program (e.g. C legacy code) from an mruby program.
The mruby-TECS bridge provides native libraries for mruby.
It also gives TECS components to receive the invocation from an mruby program.
The mruby-TECS bridge is described in more detail bellow.

As the target RTOS, TOPPERS/HRP2 \cite{url:HRP2}, \cite{par:hr-tecs} is used in this paper.
TOPPERS/HRP2 is an RTOS based on $\mu$ITRON \cite{par:microITRON} with memory protection.
However, mruby on TECS does not depend on the RTOS because TECS supports not only TOPPERS/HRP2 but also the other RTOSs such as OSEK \cite{par:OSEK} and TOPPERS/ASP \cite{url:ASP}.

\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/mrubyontecs.pdf}
    \caption{System Model of existing mruby on TECS}
    \label{fig:mrubyontecs}
\end{figure}

\subsection{mruby-TECS Bridge}\mbox{}\\

@@

Figure \ref{fig:mruby_TECS_bridge} shows an example of use of an mruby-TECS bridge for controlling a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.
\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/mruby_TECS_bridge.pdf}
    \caption{mruby-TECS Bridge}
    \label{fig:mruby_TECS_bridge}
\end{figure}

@@

\chapter{Design and Implementation}
\label{sec:Design and Implementation}
Figure \ref{fig:system_model} shows the detailed system model of the proposed framework.
Each mruby application bytecode transferred from the host is received by the loader in the RiteVM.
RiteVM reads the own bytecode and executes it.
The mruby applications run at the same time because of synchronization processing.
RiteVM scheduler switches tasks as multiple tasks can run in concurrent.

\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/system_model.pdf}
    \caption{Detailed System Model of the proposed framework}
    \label{fig:system_model}
\end{figure}

\section{mruby Bytecode Loader Using Bluetooth}
\label{sec:mruby bytecode loader using Bluetooth}
This section describes an additional feature of mruby on TECS, mruby bytecode loader using Bluetooth.
In the current system, all binary data including bytecodes are saved in an SD card.
Developers must insert and pull out the SD card every time the application programs are modified.
The OS in the target device should be also restarted.
It causes a low development efficiency to repeat that.
mruby bytecode loader using Bluetooth makes the developers' burden decrease.
%The work that an SD card is in and out needs to be done just once in the beginning.
Developers should just once carry out the work to insert the SD card and start the OS.

mruby programs are consisted of mruby application and libraries.
mruby application is the main code, which developers should program.
mruby libraries are the codes defining the functions for application such as Ruby classes. 
The mruby bytecodes including mruby application and libraries can be sent and run.
However it is wasteful in terms of the size and time to send because libraries are not frequently modified. 
In the proposed framework, mruby applications are only sent, and mruby libraries are preserved in the SD card beforehand.

The development flow in mruby bytecode loader using Bluetooth is shown in Figure \ref{fig:bluetooth_loader}.
The platform including RiteVM and mruby library is compiled and copied in the SD card at the first.
%These are assumed not to be modified.
%The binary data transferred with Bluetooth is the bytecode of the main source.
In the host, the mruby application programs (.rb) are edited and compiled into the bytecodes (.mrb) by an mruby compiler.
The generated bytecodes are transferred from the host to the target device with Bluetooth.

Moreover, developers can save the time of Bluetooth pairing since the loader can continuously load the bytecode.

\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/bluetooth_loader.pdf}
    \caption{Development Flow in mruby bytecode loader using Bluetooth}
    \label{fig:bluetooth_loader}
\end{figure}

\subsection{Component of RiteVM with mruby bytecode loader using Bluetooth} \mbox{}\\

The proposed framework provides mruby bytecode loader using Bluetooth as TECS component.
A component of mruby bytecode loader using Bluetooth is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages RiteVM configuration such as automatically generates the bytecode in the build description.
This generated bytecode is prepared beforehand in the SD card such as mruby libraries, and different from a bytecode transferred with Bluetooth.

Figure \ref{fig:component_bluetooth} shows a component diagram of MrubyTask1 and MrubyBluetooth1 {\it cell}s.
The MrubyTask1 {\it cell} is a componentized task of the RTOS (TOPPERS/HRP2).
TOPPERS/HRP2 is described in \cite{url:HRP2}, \cite{par:hr-tecs}.
The MrubyBluetooth1 is a component of RiteVM with mruby bytecode loader using Bluetooth.
A bytecode in the host is transferred and received at the top of the component.
In this framework, ZMODEM \cite{par:zmodem} is used as a binary transfer protocol.

\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/component_bluetooth.pdf}
    \caption{Component Diagram of mruby bytecode loader using Bluetooth}
    \label{fig:component_bluetooth}
\end{figure}

Figure \ref{fig:control_flow} shows the process of executing mruby program in a component of RiteVM with mruby bytecode loader using Bluetooth such as MrubyBluetooth1.
The concrete main code of tMrubyBluetooth is shown in Figure \ref{maincode_mrubybluetooth}.

First, a pointer of {\it mrb\_state} and {\it mrbc\_context} are initialized.
{\it mrb\_state} is a set of states and global variables used in mruby.

Second, RiteVM reads the bytecode of mruby libraries.
mruby libraries are a set of Ruby classes such as motor class and sensor class.
For example, motor class defines methods to rotate and stop a motor.
The tMrubyBluetooth {\it cell} has two attributes as shown in Figure \ref{celltype_mrubybluetooth}: {\it mrubyFile} and {\it irep}.
The {\it mrubyFile} indicates the program files of mruby libraries.
{\it [omit]} is only used for the TECS generator, thus the attribute, {\it mrubyFile}, does not consume memory.
The {\it irep} is the pointer of the array stored the bytecode of mruby libraries.
In short, the bytecode of mruby libraries is stored as an attribute of the component when compiling for the first time.

Third, RiteVM reads the bytecode of the mruby application transferred with Bluetooth.
The mruby application bytecode is stored in an array of type uint8\_t, which is the same as type char.
The array is different from that of holding the mruby library bytecode.
Two bytecodes are read separately in the RiteVM.

Finally, the mruby task runs.
When the mruby application is modified, only the bytecode of the modified application should be transferred.
mruby libraries need not be touched because libraries are not normally changed.
\begin{figure}[t]
    \centering
    \includegraphics[width=10cm,clip]{../EMSOFT2016/figure/control_flow.pdf}
    \caption{Control Flow of mruby bytecode loader using Bluetooth}
    \label{fig:control_flow}
\end{figure}
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tMrubyBluetooth{
    entry sTaskBody eMrubyBody;
    attr{
      [omit]char_t *mrubyFile;
      char_t *irep=C_EXP("&$cell_global$_irep");
%      uint8_t irepApp[COUNT_STK_T(TMAX_APP_BINARY_SIZE)]
%              __attribute__((nocommon));
    };
};
\end{lstlisting}
\caption{Celltype Description for mruby bytecode loader using Bluetooth}
\label{celltype_mrubybluetooth}
\end{figure}
\begin{figure}[t]
\centering
\begin{lstlisting}
void
eMrubyBody_main(CELLIDX idx)
{
  /* Declaration variables */
  mrb_state *mrb;
  mrbc_context *c;
  /* New interpreter instance */
  mrb = mrb_open();
  // Omit: error check for mrb_state
  /* New mruby context */
  c = mrbc_context_new(mrb);
  // Omit: initialization of mruby-TECS bridge
  /* Receive the bytecode via Bluetooth */
  bluetooth_loader( App_bytecode );
  /* Load mruby library bytecode and run */
  mrb_load_irep_cxt(mrb, ATTR_irep, c);
  /* Load mruby application bytecode and run */
  mrb_load_irep_cxt(mrb, App_bytecode, c);
  if (mrb->exc) {
    /* Failure to execute */
    mrb_p(mrb, mrb_obj_value(mrb->exc));
    exit(0);
  }
  /* Free mruby context */
  mrbc_context_free(mrb, c);
  /* Free interpreter instance */
  mrb_close(mrb);
}

@@

One of approaches for multitasking is co-routine.
Co-routine is a cooperative thread, and scheduled by developers with the functions such as {\it resume} and {\it yield}. 
(Ruby co-routine is defined in class Fiber \cite{url:co-routine})
Co-routine is a non-preemptive multitasking, and does not receive the OS's support because developers have to switch tasks manually.
Co-routine can not take advantage of multi core processing.

Besides, as another method, {\it delay}, a service call of $\mu$ITRON, can be used for multitasking.
This service call delays the execution of the own task for the time of the argument.
In the approach of {\it delay} function, developers have to schedule as the same as co-routine.

In the proposed framework, as an approach for multitask processing, RiteVM scheduler is implemented.
The RiteVM scheduler is a fair scheduler and runs multiple tasks equally.

\subsection{RiteVM Scheduler} \mbox{}\\

The RiteVM scheduler is a cyclic handler, and calls {\it rotateReadyQueue}, a service call of $\mu$ITRON to switch tasks with the same priority.
In other words, the RiteVM scheduler calls {\it rotateReadyQueue} cyclically.
{\it rotateReadyQueue} is described as follows.

The case is assumed that there are two tasks with the same priority, and both tasks are in an infinite loop.
In the current system, when one task is executed first, another task would not be executed.
That is because the task with first execution runs in the loop.
